---
title: "Final Project: Insurance"
author: "Rebecca Nguyen"
output:
  pdf_document: default
---
The purpose of this final project is to explore the health insurance data set 
and predict insurance charges based on several determinants.

# Load appropriate packages/csv file
```{r}
library(tidyverse)
insurance_df <- read_csv('insurance.csv')
```

## Data Exploration
```{r}
head(insurance_df)
str(insurance_df)
```
1338 observations  
7 variables:   
numerical: age, bmi, children, charges | character: sex, smoker, region
     
```{r}
table(insurance_df$sex) 
gender_num <- data.frame(value = c(662, 676),
                        Gender = c("females", "males"))
ggplot(gender_num, aes(x = "", y = value, fill = Gender)) +
  geom_col(color = "black") +
  geom_text(aes(label = value),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_brewer() +
  theme_void()
```

```{r}
insurance_df %>%
  group_by(sex) %>%
  summarize(avg = mean(age, na.rm = TRUE), sd = sd(age, na.rm = TRUE))
```

```{r}
table(insurance_df$children) 
children_num <- data.frame(value = c(574, 324, 240, 157, 25, 18),
                           Num_children = c("None", 1, 2, 3, 4, 5))
ggplot(children_num, aes(x = "", y = value, fill = Num_children)) +
  geom_col(color = "black") +
  geom_text(aes(label = value),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_brewer() +
  theme_void()
```

```{r}
table(insurance_df$smoker) 
smoke_num <- data.frame(value = c(1064, 274),
                        Smoker = c("No", "Yes"))
ggplot(smoke_num, aes(x = "", y = value, fill = Smoker)) +
  geom_col(color = "black") +
  geom_text(aes(label = value),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_brewer() +
  theme_void()
```

```{r}
table(insurance_df$region)
region_num <- data.frame(value = c(324, 325, 364, 325),
                         Region = c("Northeast", "Northwest", "Southeast",
                                    "Southwest"))
ggplot(region_num, aes(x = "", y = value, fill = Region)) +
  geom_col(color = "black") +
  geom_text(aes(label = value),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_brewer() +
  theme_void()
```

```{r}
mean(insurance_df$charges)
```
The overall average insurance charge is $13,270.42

# Sex vs. Charges Significance
```{r}
sex_plot <- insurance_df %>% 
  ggplot(aes(x = sex, y = charges, color = sex)) +
  facet_wrap(~ region) +
  geom_boxplot(outlier.alpha = 0.5)
sex_plot
sex_sum <- insurance_df %>%
  group_by(sex) %>%
  summarize(avg = mean(charges, na.rm = TRUE), sd = sd(charges, na.rm = TRUE))
sex_sum 
```
This histogram illustrates how there is no apparent bias between gender 
and insurance charges. Regardless of gender, the mean insurance charges are 
relatively similar as are the median charges per region.

# Age vs. Charges Significance
```{r}
age_plot <- insurance_df %>% 
  ggplot(aes(x = age, y = charges, color = smoker)) + 
  facet_wrap(~ region) +
  geom_point()
age_plot
cor(insurance_df$age, insurance_df$charges) 
```
Based on the graphs, there seems to be a positive, linear relationship between 
age and insurance charges. As age increases, so does the health insurance. 
There also seems to be 2 distinct levels above the baseline that indicates
being a smoker results in elevated charges.
However, the correlation coefficient indicates a weak to moderate linear 
relation between age and insurance charges.

# BMI vs. Charges Significance
```{r}
BMI_plot <- insurance_df %>% 
  ggplot(aes(x = bmi, y = charges)) + 
  facet_wrap(~ region) +
  geom_point()
BMI_plot
cor(insurance_df$bmi, insurance_df$charges)
```
Based on the graphs and correlation coefficient, there appears to be a weak, 
positive linear relationship between BMI and insurance charges. However, NE & 
NW look similar to each other as well as SE & SW.

# Children vs. Charges Significance
```{r}
child_plot <- insurance_df %>%
  ggplot(aes(x = children, y = charges, fill = children, group = children)) +
  geom_boxplot(outlier.alpha = 0.5)
child_plot
cor(insurance_df$children, insurance_df$charges)
```
Based on the graph & correlation coefficient, there is no correlation between 
the number of offsprings & insurance charges.

# Smoker vs. Charges Significance
```{r}
smoke_plot <- insurance_df %>%
  ggplot(aes(x = smoker, y = charges, color = smoker)) +
  geom_boxplot(outlier.alpha = 0.5)
smoke_plot
smoke_sum <- insurance_df %>%
  group_by(smoker) %>%
  summarize(avg = mean(charges, na.rm = TRUE), sd = sd(charges, na.rm = TRUE))
smoke_sum
```
This graph illustrates a startling, significant difference between smoker 
status and health insurance charges. A smoker is expected to pay on average
$32,050, whereas a non-smoker is expected to pay on average $8,434. That is a
$23,616 difference!
```{r}
smokebmi_plot <- insurance_df %>%
  ggplot(aes(x = bmi, y = charges, color = smoker)) +
  geom_point()
smokebmi_plot
insurance_df %>% 
  ggplot(aes(x = bmi, y = charges, color = smoker)) + 
  facet_wrap(~ region) +
  geom_point()
```
This graph indicates a positive, moderate linear relation between being a 
smoker as well as obese (> 30 BMI) and health insurance charges. Health
insurance increases as the BMI of a smoker increases. It is apparent that
the base charges of smokers are higher than their counterparts. It is rather evident when visualizing per region.
```{r}
bmismc <- insurance_df %>%
  filter(smoker == "yes") %>%
  select(bmi, charges)
bmismc
lm(charges~bmi, bmismc)
lm_bmismcplot <- bmismc %>%
  ggplot(aes(bmi, charges)) +
  geom_point() +
  geom_abline(intercept = -13187, slope = 1473)
lm_bmismcplot
```
As a smoker by default, one unit increase in BMI results in 1473 increase in
charges.

# Region vs. Charges Significance
```{r}
region_plot <- insurance_df %>% 
  ggplot(aes(x = region, y = charges)) + 
  geom_boxplot(outlier.alpha = 0.5)
region_plot
region_sum <- insurance_df %>%
  group_by(region) %>%
  summarize(avg = mean(charges, na.rm = TRUE), sd = sd(charges, na.rm = TRUE))
region_sum
```
Based on the box plot and summary, there is very weak relation between region 
and insurance. Regardless of location, the mean & median insurance charges are relatively similar. However, the east seems to charge a bit more than the west.
The southeast being the most expensive.

**From this data exploration, we can conclude that age and BMI alongside smoker
play a significant role in health insurance charges. Sex, children, and region 
do not significantly affect health insurance charges.** 

# Linear Regression Model
```{r}
linear_model_insurance <- lm(charges~., insurance_df)
linear_model_insurance
summary(linear_model_insurance)
```
Based on the summary, the most significant factors are age, BMI, children, &
smoker. The least significant is the regions. Sex holds no significance in relation to charges. 
The p-value for sexmale is big, so let's take it out
```{r}
linear_model_insurance2 <- lm(charges~.-sex, insurance_df)
linear_model_insurance2
summary(linear_model_insurance2)
```
The multiple R-squared:  0.7509 tells us this linear regression model decently
explains the data.

```{r}
lm_test <- lm(charges ~.+bmi*smoker, insurance_df)
summary(lm_test)
```
By including the synergistic effect of BMI & smoker to LM, the new multiple 
R-squared increased to 0.840, which directly translates to a better model.
```{r}
lm_predict <- predict(lm_test, insurance_df)
test <- data.frame("Predicted" = lm_predict, "Actual" = insurance_df$charges)
ggplot(test, aes(Actual, Predicted)) + geom_point() + geom_abline()
```
After adjusting the LM, we can see the model being the best fit for the
data.

# Prediction Application
```{r}
age = 60
sex = "female"
bmi = 35
children = 2
smoker = "yes"
region = "northeast"
single_obs <- data.frame(age = 60,
                         sex = "female",
                         bmi = 35,
                         children = 2,
                         smoker = "yes",
                         region = "northeast")
single_obs
predict(lm_test, single_obs)
```
$45,503.18  is our prediction.