---
title: "Health Insurance: Data Exploration & Linear Model"
author: "Rebecca Nguyen"
output:
  pdf_document: default
---
The purpose of this project is to explore the health insurance data set 
and predict insurance charges based on several determinants.  

Healthcare should be easily accessible and affordable for all regardless of our medical history. It is a fundamental right that everyone should have. We see many countries offering free, universal healthcare, such as Mexico, Canada, and the UK. Yet, the US does not offer this form of healthcare, even though we are a highly-developed country.  

That being said, we all know about that rumor that US health insurance prices are based off of the presence of medical conditions, or drug use, to name a couple. But is it actually true? Are these health insurance companies deliberately increasing prices and making money off of us for not being "healthy"? What significant factors greatly affect how much we are charged for health insurance? This is what I would like to explore through this data set. 

# Load appropriate packages/csv file
```{r}
library(tidyverse)
insurance_df <- read_csv('insurance.csv')
```

## Data Exploration
```{r}
head(insurance_df)
str(insurance_df)
```
1338 observations  
7 variables:   
numerical: age, bmi, children, charges | character: sex, smoker, region
     
```{r}
table(insurance_df$sex) 
gender_num <- data.frame(value = c(662, 676),
                        Gender = c("females", "males"))
ggplot(gender_num, aes(x = "", y = value, fill = Gender)) +
  geom_col(color = "black") +
  geom_text(aes(label = value),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_brewer() +
  theme_void()
```

```{r}
insurance_df %>%
  group_by(sex) %>%
  summarize(avg = mean(age, na.rm = TRUE), sd = sd(age, na.rm = TRUE))
```

```{r}
table(insurance_df$children) 
children_num <- data.frame(value = c(574, 324, 240, 157, 25, 18),
                           Num_children = c("None", 1, 2, 3, 4, 5))
ggplot(children_num, aes(x = "", y = value, fill = Num_children)) +
  geom_col(color = "black") +
  geom_text(aes(label = value),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_brewer() +
  theme_void()
```

```{r}
table(insurance_df$smoker) 
smoke_num <- data.frame(value = c(1064, 274),
                        Smoker = c("No", "Yes"))
ggplot(smoke_num, aes(x = "", y = value, fill = Smoker)) +
  geom_col(color = "black") +
  geom_text(aes(label = value),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_brewer() +
  theme_void()
```

```{r}
table(insurance_df$region)
region_num <- data.frame(value = c(324, 325, 364, 325),
                         Region = c("Northeast", "Northwest", "Southeast",
                                    "Southwest"))
ggplot(region_num, aes(x = "", y = value, fill = Region)) +
  geom_col(color = "black") +
  geom_text(aes(label = value),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  scale_fill_brewer() +
  theme_void()
```

```{r}
mean(insurance_df$charges)
```
The overall average insurance charge is $13,270.42

# Sex vs. Charges Significance
```{r}
sex_plot <- insurance_df %>% 
  ggplot(aes(x = sex, y = charges, color = sex)) +
  facet_wrap(~ region) +
  geom_boxplot(outlier.alpha = 0.5)
sex_plot
sex_sum <- insurance_df %>%
  group_by(sex) %>%
  summarize(avg = mean(charges, na.rm = TRUE), sd = sd(charges, na.rm = TRUE))
sex_sum 
```
The boxplots illustrate how there is no apparent bias between gender 
and insurance charges. Regardless of gender, the mean insurance charges are 
relatively similar as are the median charges per region.

# Age vs. Charges Significance
```{r}
age_plot <- insurance_df %>% 
  ggplot(aes(x = age, y = charges, color = smoker)) + 
  facet_wrap(~ region) +
  geom_point()
age_plot
cor(insurance_df$age, insurance_df$charges) 
age_smoker <- insurance_df %>%
  filter(smoker == "yes")
age_nosmoker <- insurance_df %>%
  filter(smoker == "no")
cor(age_smoker$age, age_smoker$charges)
cor(age_nosmoker$age, age_nosmoker$charges)
```
Based on the scatter plots, there seem to be a positive, linear relationship between 
age and insurance charges. As age increases, so does the health insurance. 
There also seems to be 2 distinct levels above the baseline that indicates
being a smoker results in elevated charges. Other unexplored factors could also contribute to this phenomenon, such as the type of insurance.  

However, the correlation coefficient indicates a weak to moderate linear 
relation between age and insurance charges. When comparing age, non-smoker, and charges, the correlation coefficient indicates a strong linear relationship. When comparing age, smoker, and charges, the correlation coefficient indicates a weak to moderate linear relationship.

# BMI vs. Charges Significance
```{r}
BMI_plot <- insurance_df %>% 
  ggplot(aes(x = bmi, y = charges)) + 
  facet_wrap(~ region) +
  geom_point()
BMI_plot
cor(insurance_df$bmi, insurance_df$charges)
```
Based on the scatter plots and correlation coefficient, there appears to be a weak, 
positive linear relationship between BMI and insurance charges. However, NE & 
NW look similar to each other as well as SE & SW.

# Children vs. Charges Significance
```{r}
child_plot <- insurance_df %>%
  ggplot(aes(x = children, y = charges, fill = children, group = children)) +
  geom_boxplot(outlier.alpha = 0.5)
child_plot
cor(insurance_df$children, insurance_df$charges)
```
Based on the box plot & correlation coefficient, there is no correlation between 
the number of offsprings & insurance charges.

# Smoker vs. Charges Significance
```{r}
smoke_plot <- insurance_df %>%
  ggplot(aes(x = smoker, y = charges, color = smoker)) +
  geom_boxplot(outlier.alpha = 0.5)
smoke_plot
smoke_sum <- insurance_df %>%
  group_by(smoker) %>%
  summarize(avg = mean(charges, na.rm = TRUE), sd = sd(charges, na.rm = TRUE))
smoke_sum
```
This box plot illustrates a startling, significant difference between smoker 
status and health insurance charges. A smoker is expected to pay on average
$32,050, whereas a non-smoker is expected to pay on average $8,434. That is a
$23,616 difference!
```{r}
smokebmi_plot <- insurance_df %>%
  ggplot(aes(x = bmi, y = charges, color = smoker)) +
  geom_point()
smokebmi_plot
insurance_df %>% 
  ggplot(aes(x = bmi, y = charges, color = smoker)) + 
  facet_wrap(~ region) +
  geom_point()
```
These scatter plots indicate a positive, moderate linear relation between being a 
smoker as well as obese (> 30 BMI) and health insurance charges. Health
insurance increases as the BMI of a smoker increases. It is apparent that
the base charges of smokers are higher than their counterparts. It is rather evident when visualizing per region.
```{r}
bmismc <- insurance_df %>%
  filter(smoker == "yes") %>%
  select(bmi, charges)
bmismc
lm(charges~bmi, bmismc)
lm_bmismcplot <- bmismc %>%
  ggplot(aes(bmi, charges)) +
  geom_point() +
  geom_abline(intercept = -13187, slope = 1473)
lm_bmismcplot
```
As a smoker by default, one unit increase in BMI results in 1473 increase in
charges.

# Region vs. Charges Significance
```{r}
region_plot <- insurance_df %>% 
  ggplot(aes(x = region, y = charges)) + 
  geom_boxplot(outlier.alpha = 0.5)
region_plot
region_sum <- insurance_df %>%
  group_by(region) %>%
  summarize(avg = mean(charges, na.rm = TRUE), sd = sd(charges, na.rm = TRUE))
region_sum
```
Based on the box plot and summary, there is very weak relation between region 
and insurance. Regardless of location, the mean & median insurance charges are relatively similar. However, the east seems to charge a bit more than the west.
The southeast being the most expensive.

**From this data exploration, we can conclude that age and BMI alongside smoker
play a significant role in health insurance charges. Sex, children, and region 
do not significantly affect health insurance charges.** 

# Linear Regression Model
```{r}
linear_model_insurance <- lm(charges~., insurance_df)
linear_model_insurance
summary(linear_model_insurance)
```
Based on the summary, the most significant factors are age, BMI, children, &
smoker. The least significant is the regions. Sex holds no significance in relation to charges. 
The p-value for sexmale is big, so let's take it out.
```{r}
linear_model_insurance2 <- lm(charges~.-sex, insurance_df)
linear_model_insurance2
summary(linear_model_insurance2)
```
The multiple R-squared:  0.7509 tells us this linear regression model decently
explains the data.

```{r}
lm_test <- lm(charges ~.+bmi*smoker, insurance_df)
summary(lm_test)
```
By including the synergistic effect of BMI & smoker to LM, the new multiple 
R-squared increased to 0.840, which directly translates to a better model.
```{r}
lm_predict <- predict(lm_test, insurance_df)
test <- data.frame("Predicted" = lm_predict, "Actual" = insurance_df$charges)
ggplot(test, aes(Actual, Predicted)) + geom_point() + geom_abline()
```
After adjusting the LM, we can see the model being the best fit for the
data.

# Prediction Application
```{r}
age = 60
sex = "female"
bmi = 35
children = 2
smoker = "yes"
region = "northeast"
single_obs <- data.frame(age = 60,
                         sex = "female",
                         bmi = 35,
                         children = 2,
                         smoker = "yes",
                         region = "northeast")
single_obs
predict(lm_test, single_obs)
```
$45,503.18  is our prediction.

# Challenges
A significant challenge that I faced while working on this project was the data visualization interpretation. I had difficulty analyzing the plots I created, such as identifying patterns in the graphs and determining what factors may have caused a certain graph to appear that way. For example, initially, I colored the data points of the "Age vs. Charges Significance" graph to show the gender of each point. When looking at it, I did not understand why there were various levels above the apparent baseline. I tinkered with the color feature and found coloring the data points to show smoker status to really explain why the points were the way they were. Being a smoker results in a significant increase in charges throughout life. Other unexplored variables could also explain why there are more levels to the graph, such as the type of insurance. 